TAPIR: Threaded Application Programming on Intermittent Resources

### 2017-05-11 paper.tex
introduction first draft

\section{Introduction/Background} -Amanda
  \label{sec:intro}
  Chain, DINO\\
  Multithreading\\
  Intermittance Review\\
  RMW\\
  Example Citation\cite{Chain, Mementos, Aware, Dino}!

@@

\section{Related Work} -Emily
Operating system work\\
TU delft guys...\\
Energy Align\\
TinyOS stuff in wireless sensor networks\\

>>>>

\section{Introduction/Background}
  \label{sec:intro}
%A/N: I don't think that the paragraphs in this section are in a great order, but I am having a hard time making it flow better.
%Also I'm absolutely sure each of these paragraphs can definitely be bulked up more for length. Please help thank you!!!
%Also also: I'm trying to get [energy-harvesting computing devices] and [intermittent execution model] below to be in italics,
%but when I use \textit it makes the ENTIRE PAPER in italics, i used the curly braces and everything helpppppp that has never happened to me before

The increase in popularity of Internet of Things (IoT) has led to research focus in wireless technologies and low-power computing.
New technologies in both of these areas has led to new [energy-harvesting computing devices] that operates only with power it has 
collected from the surrounding environment \cite{Chain}. These devices have use cases in medical, IoT, and research applications.

Commonly, an energy-harvesting devices charges and collects power in a storage capacitor until the capacitor reaches a sufficient level of charge.
Once the capacitor is charged, the device begins its operation phase and will run until the capacitor is drained.
This leads to the [intermittent execution model] where regular shut downs and reboots to recharge are the common case rather than the rare case in
a traditional execution model. Because the intermittent execution model includes power failures and reboots, continuous execution programs behave
irregular behavior on an intermittent device and a new programming model is needed.

Our multithreaded model is build on two existing programming models, DINO and Chain.
An initial intermittent programming model DINO breaks intermittent programs into tasks, which creates automatic checkpointing locations 
during exeuction \cite{Dino}. DINO uses a checkpointing and recovery model that tracks both volatile and nonvolatile memory.
The DINO execution model executes all tasks transactionally, leading to atomic task semantics and does not permit consistency violations in 
nonvolatile state \cite{Dino}. Chain, a more recent iteration of the concepts in DINO, uses volatile state rarely, if ever, and uses task boundaries
called channels to store most used data in nonvolatile memory \cite{Chain}. At compile time, Chain allocates the necessary channels between tasks
by parsing channel read and write statements in order to match up aligning tasks. All inputs and outputs to each task, as well as necessary data
structures, are stored in a channel for a respective task. A channel may only be modified by one task, and only one task may run at a time,
which mitigates time of check, time of use bugs (TOCTOU) in intermittence. 
The one exception is a self-channel for recursive code, which is double buffered.
Keeping all modifiable data in nonvolatile memory that is not open to easy rewrite makes these tasks idempotent across reboots.
Because of continuous reboots, many operations necessarily will be executed more than once,
and correctness is compromised if those operations are not guaranteed to have the same outcome by the programming model after each reboot.

A main issue in intermittent computing along with idempotence is the Read-Modify-Write (RMW) problem. 
Because volatile memory is cleared on each reboot, it is possible that in between reading and modifying or modifying and writing the data had changed,
leading to incorrect answers and an inconsistent state. In the Chain programming model, this problem is solved by creating a new task to house the
problematic operations. Because tasks are blocked by channels that can only be written by one task and are saved to nonvolatile memory, RMW is avoided.
However, in multiprogramming the problem manifests in another way; if the scheduler switches tasks on a reboot, the currently running thread could have
been in the middle of its critical section, making it difficult for other threads to make progress if context switches always occur after a reboot.
Because of this we have decided on a round-robin scheduler with context switches at the end of a task, to guarantee forward progress while knowing
which tasks have run to completion.

In the current Chain programming model, writing code that completes multiple objectives correctly is now difficult. 
Current chain programs are small and only have one objective. Additionally, there is no support for advanced control flow.
Multiple control threads may exist, but in the current model there is no support for them and programmability is difficult and often incorrect.
Supporting multiple control threads in Chain also necessitates a linear increase in a limited memory resource, because all large shared data structures
must be copied into each necessary channel, and also makes updating the structures a time consuming, high-overhead task.
In order to increase functionality of chain and to mitigate these shortcomings, we add familiar threading semantics to Chain in order to express
multi-threaded control flow operation. In order to do this we create a lightweight scheduler task that becomes part of the inter-task checkpointing
operation and provide a method of sharing large data structures between tasks with a locking feature. This method is low-memory and low-execution
overhead, which can then maintain the Chain programming semantics at the user level still successfully complete work on an intermittent device.

Example Citation\cite{Chain, Mementos, Aware, Dino}!

@@

\section{Related Work} -Emily
Operating system work\\
TU delft guys...\\
Energy Align\\
